Elimination Of Jump Objects

### 2022-06-06 paper/skeleton.tex

\title{\textbf{Elimination Of Jump Objects}}
\author{\textbf{author}}
\date{May 2022}

@@

%%We also assume that all jumps are conditional. If this is not the case, then we can easily convert an unconditional goto to a conditional goto by adding an additional condition: \emph{if (TRUE)}.

@@

\subsubsection{Goto forward wrapped in an object}
So, after the conversion described in section 3.1.1, in general cases, forward jumps will look like this:
\begin{ffcode}
goto
  [g]
    |\sgm{1}|
      |\sgm{2}|
      if.
        |\m{1}|
        g.forward |\dlt{1}|
        |\dlt{2}|
      |\sgm{3}|
\end{ffcode}

Let's denote lines 20-26 as X.

@@

Also, after replacing the jump, we must create an additional \emph{flag} object, which will change if the jump condition is met, i.e. in the new if-statement, the 'ELSE' branch will be executed. We need this so that when a series of nested objects are nested one in another, when the jump condition is met, subsequent objects whose nesting is less than the nesting of the original object containing goto forward and more than the abstract object \emph{g} and following after the original object are not executed.

@@

Thus, after transformation, the program will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}|
  |\sgm{2}|
    if.
      not. (|\m{1}|)
      seq
        |\dlt{2}|
        |\sgm{3}|
      seq
        |\dlt{1}|
        flag.write 1
\end{ffcode}

Let's denote lines 27-34 as X'.

@@

In general, using goto forward inside a while-loop will look like this:
\begin{ffcode}
goto
  [g]
    |\sgm{0}|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.forward |\dlt{1}|
            |\dlt{2}|
          |\sgm{3}|
\end{ffcode}

Let's denote lines 37-46 as Y.

@@

Thus, transformation will look like this:
\begin{ffcode}
memory > flag
flag.write 0
while.
  and.
    not. (eq. (flag 1))
    |\m{0}|
  |\sgm{1}|
    |\sgm{2}|
    if.
      not. (|\m{1}|)
      seq
        |\dlt{2}|
        |\sgm{3}|
      seq
        |\dlt{1}|
        flag.write 1
\end{ffcode}

Let's denote lines 49-62 as Y'.

@@

After transformation it will look like:
\begin{ffcode}
memory > flag
flag.write 0
while.
  and.
    not. (eq. (flag 1))
    |\m{9}|
  |\sgm{9}|
    Y'
    if.
      not. (eq. (flag 1))
      |\sgm{4}|
\end{ffcode}

\subsection{backward jump}
Backward jump in EO is implemented as:

@@

goto
 [g]
  seq
   stdout "Will be printed forever"
   g.backward
\end{ffcode}

Backward jump works the same as do-while construction in other languages: the first iteration of the loop is always executed, and the next only if the condition is TRUE.
So, all statements that met before the "jump" itself should be executed before the new while loop once, and all subsequent times were executed depending on the condition of the loop (Look at Image2).

@@

Consider the following code:
\begin{ffcode}
[] > f
 memory > i
 seq > @
  i.write 1
  goto
   [g]
    seq > @
     i.write (i.add 1)
     if.
      i.less 10
      g.backward
      TRUE
 stdout "Finished!"
\end{ffcode}

Transformation is to replace a piece of code with a while loop with a backward jump condition.

Equivalent replacement of the example described above:
\begin{ffcode}
[] > f
 memory > i
 seq > @
  i.write 1
  i.write (i.add 1)
  while.
   i.less 10
   seq
    i.write (i.add 1)
 stdout "Finished!"
\end{ffcode}

In cases where backward jump is used to implement continue, the transformation will look slightly different.

\subsubsection{continue}
Backward jump is used to implement continue statement:

\begin{ffcode}
[]
 memory > i
 goto > @
  [g]
   seq > @
    while.
     i.less 10
     seq
      i.write (i.add 1)
      if.
       eq. (i.mod 5 0)
       g.backward
       TRUE
      stdout "Not divisible by 5!\n"
\end{ffcode}

This statement is logically the same as a break statement, except that after a break is executed the entire while loop stops. When you execute the continue statement, the next iteration of the loop is simply started.
So from a logical standpoint, replacing continue statement is identical to replacing break statement described in section 3.1.2, except that now we don't have to create an auxiliary variable to abort the execution of the while loop. In this case, the transformation consists in inverting the jump condition and executing the remaining statements in the body of the new if statement.

Modification of the given example code will look like:
\begin{ffcode}
[]
 memory > i
 seq > @
 while.
  i.less 10
  seq
   i.write (i.add 1)
   if.
    not. (eq. (i.mod 5 0))
    stdout "Not divisible by 5!\n"
    TRUE
\end{ffcode}

\subsection{Complex cases}
In complex cases, when one goto object is nested in another, the transformations will be similar to cases, described in sections 3.1 - 3.2. When Backward jump, we should insert a while loop into the body all the statements that stand between the declaration of the object and the jump itself, as described at Section 3.2 . When Forward jump, remember all statements that follow in the body of if-statement where the jump is made and execute these statements after the GOTO object, as decribed at Section 3.1 .

Consider the following complex example with nested goto-objects:
\begin{ffcode}

@@

Considering all the jumps that appear in the code (the jumps themselves, not the declaration of goto objects), we apply the transformations already discussed and proved by us above one by one.

Applying all the transformations to each encountered goto object, we get the following code:
\begin{ffcode}

@@

\subsection{Final algorithm}
Summarizing the points above, we can fully describe the algorithm itself. It consists of three steps (See Image3). The first step is to convert all unconditional goto objects to conditional goto objects. In the second step, applying the transformations described above, we replace jumps with semantically equivalent pieces of code. In the third step, remove all the declarations of the GOTO objects themselves.

>>>>

\title{Elimination Of Jump Objects}
\author{author}

@@

@@

\subsubsection{Goto forward wrapped in an object}
So, after the conversion described in section 3.1.1, in general cases, forward jumps will look like this:
\begin{ffcode}
goto
  [g]
    |\sgm{1}| |$\label{ln:ret0}$|
      |\sgm{2}|
      if.
        |\m{1}|
        g.forward |\dlt{1}|
        |\dlt{2}|
      |\sgm{3}| |$\label{ln:ret1}$|
\end{ffcode}

Let's denote lines \ref{ln:ret0}-\ref{ln:ret1} as X.

@@

Also, after replacing the jump, we must create an additional \emph{flag} object, which will change if the jump condition is met, i.e. in the new if-statement, the 'ELSE' branch will be executed. We need this so that when a series of nested objects are nested one in another, when the jump condition is met, subsequent objects whose nesting is less than the nesting of the original object containing goto forward and more than the nesting of an abstract object \emph{g} and following after the original object are not executed.

@@

Thus, after transformation, the program will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}||$\label{ln:ret2}$|
  |\sgm{2}|
  if.
    |\m{1}|.not
    |\dlt{2}|
    seq
      |\dlt{1}|
      flag.write 1
  if.
    (eq. (flag 1)).not
    |\sgm{3}|
    TRUE|$\label{ln:ret3}$|
\end{ffcode}

Let's denote lines \ref{ln:ret2}-\ref{ln:ret3} as X'.

Now let's look at an example program with nested objects containing goto forward:
\begin{ffcode}
goto
  [g]
    |\sgm{0}|
      X
      |\sgm{4}|
\end{ffcode}

Applying the transformation, we get:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}|
  X'
  if.
    (eq. (flag 1)).not
    |\sgm{4}|
    TRUE
\end{ffcode}

@@

In general, using goto forward inside a while-loop will look like this:
\begin{ffcode}
goto
  [g]
    |\sgm{0}||$\label{ln:ret4}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.forward |\dlt{1}|
            |\dlt{2}|
          |\sgm{3}||$\label{ln:ret5}$|
\end{ffcode}

Let's denote lines \ref{ln:ret4}-\ref{ln:ret5} as Y.

@@

Thus, transformation will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}||$\label{ln:ret6}$|
  while.
    and.
      (eq. (flag 1)).not
      |\m{0}|
    |\sgm{1}|
      |\sgm{2}|
      if.
        |\m{1}|.not
        |\dlt{2}|
        seq
          |\dlt{1}|
          flag.write 1
      if.
        (eq. (flag 1)).not
        |\sgm{3}|
        TRUE|$\label{ln:ret7}$|
\end{ffcode}

Let's denote lines \ref{ln:ret6}-\ref{ln:ret7} as Y'.

@@

After transformation it will look like:
\begin{ffcode}
memory > flag
flag.write 0
while.
  and.
    (eq. (flag 1)).not
    |\m{9}|
  |\sgm{9}|
    Y'
    if.
      (eq. (flag 1)).not
      |\sgm{4}|
      TRUE
\end{ffcode}

\subsection{Backward jump}
Goto backward in EO is implemented as:

@@

goto
  [g]
    |\sgm{1}|
      stdout "Will be printed forever"
      g.backward
\end{ffcode}

Same as with goto forward, there is the goto atom-object, the abstract object \emph{g}, and the g.backward attribute. Several abstract objects such as \emph{g} can be declared inside of goto-object, and the attribute \emph{.backward} can be called to each of them.

\subsubsection{Simple Backward} The simplest example of using goto-backward:

\begin{ffcode}
goto
  [g]
    g.backward
\end{ffcode}

This is identical to that discussed in Section 3.1.1. In cases where the jump is performed without condition, you need to add an additional if-statement, similar to the example in Section 3.1.1 but with goto-backward:

\begin{ffcode}
goto
  [g]
    if.
      TRUE
      g.backward
      TRUE
\end{ffcode}

\subsubsection{Goto backward wrapped in an object}
After the conversion decribed in Section 3.2.1, in general cases, backward jumps will look like this:

\begin{ffcode}
goto
  [g]
    |\sgm{1}||$\label{ln:ret8}$|
      |\sgm{2}|
      if.
        |\m{1}|
        g.backward
        |\dlt{1}|
      |\sgm{3}||$\label{ln:ret9}$|
\end{ffcode}

Let's denote lines \ref{ln:ret8}-\ref{ln:ret9} as W.

Note that goto-backward works the same as do-while construction in other languages: the first iteration of the new loop is always executed, and the next only if the condition is TRUE.
So, all objects that occurs between the "jump" itself and the declaration of an abstract object \emph{g} should be executed before the new while loop once, and all subsequent times were executed depending on the condition of the loop (Look at Image2).

As soon as the execution of the program reaches the jump condition, the first iteration of the new do-while loop will already be executed.
After that, if the jump conditions are met, all objects that are between the declaration of the abstract object \emph{g} and the jump itself will be executed again. And that means that we can create a while loop at the point of the jump, so that every time we check the condition, we don't make a jump, but just start a new iteration of the loop.

Inside the body of the new while loop, we copy all objects that are inside the abstract object g - with nesting at least the object g itself, except for the call of the jump itself with its condition.
At the point of the jump, we should create an if-statement with an inverted jump condition: in the 'THEN' branch there will be objects that were returned in the 'ELSE' branch of the original if for the jump; in the 'THEN' branch, the value of the flag object will be changed, which will mean that the condition for the jump is met, and all subsequent objects with nesting at least the nesting of the added while loop will not be executed.
To do this, for all these objects, going down the nesting from the object where the jump was made to the while loop, we need to add an if-statement that will check whether the condition for the jump has been met before.

This transformation does not violate the program logic, because if the jump condition is fulfilled, then all the objects going after the jump itself will not be executed due to the added if-statements.

@@

So, after transformation, the program will look like this:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{1}||$\label{ln:ret10}$|
  |\sgm{2}|
  if.
    |\m{1}|.not
    |\dlt{1}|
    while.
      |\m{1}|
      seq
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|.not
            |\dlt{1}|
            flag.write 1
          if.
            (eq. (flag 1)).not
            |\sgm{3}|
            TRUE
        flag.write 0
  |\sgm{3}||$\label{ln:ret11}$|

@@

Let's denote lines \ref{ln:ret10}-\ref{ln:ret11} as W'.

Let's look at an example of a chain of nested objects containing a goto backward:

\begin{ffcode}
goto
  [g]
    |\sgm{0}|
      W
      |\sgm{4}|
\end{ffcode}

Applying transformation, we get:
\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}|
  W'
  if.
    (eq. (flag 1)).not
    |\sgm{4}|
    TRUE
\end{ffcode}

\subsubsection{Goto backward inside the while-loop}
Goto backward, like goto forward, can be called inside a while-loop body.
As discussed in section 3.1.3, the behavior of while is different from other objects in the language. Therefore, the transformation will be slightly different from that described in section 3.2.2.

In general cases, a goto backward call inside a while loop will look like this:

\begin{ffcode}
goto
  [g]
    |\sgm{0}||$\label{ln:ret12}$|
      while.
        |\m{0}|
        |\sgm{1}|
          |\sgm{2}|
          if.
            |\m{1}|
            g.backward
            |\dlt{1}|
          |\sgm{3}||$\label{ln:ret13}$|

@@

Let's denote lines \ref{ln:ret12}-\ref{ln:ret13} as Z.

As described in section 3.2.2, the program runs from top to bottom until it meets goto forward and if the jump condition is met, the abstract object g is executed again.
Similar to what was said in section 3.2.2, all objects inside the abstract object g should be copied at the jump point.

The only difference, when the jump itself frames a while loop, is that when the jump condition is met, we need to interrupt the execution of all the internal while loops inside the new while loop, which we replaced the jump with.

In order to interrupt the execution of each of the internal while loops, we can add an additional condition to each of them, whether the jump condition is met or not. This can be done with an additional flag object that will signal for each inner while loop whether the jump condition has been met.

So, all internal while loops will not be executed further if flag shows that the jump condition is met and the logic of the program is not violated. For all other objects inside our new while loop, we should apply the transformation described in section 3.2.2.

The transformation will look like this:

\begin{ffcode}
memory > flag
flag.write 0
|\sgm{0}||$\label{ln:ret14}$|
  while.
    |\m{0}|
    |\sgm{1}|
      |\sgm{2}|
      while.
        |\m{1}|
        |\sgm{0}|
          while.
            and.
              (eq. (flag 1)).not
              |\m{0}|
            |\sgm{1}|
              |\sgm{2}|
              if.
                |\m{1}|.not
                |\dlt{1}|
                flag.write 1
              if.
                (eq. (flag 1)).not
                |\sgm{3}|
                TRUE
      |\sgm{3}||$\label{ln:ret15}$|

Let's denote lines \ref{ln:ret14}-\ref{ln:ret15} as Z'.

Let's take a look on a program with nested while-loops:
\begin{ffcode}
goto
  [g]
    while.
      |\m{9}|
      |\sgm{9}|
        Z
        |\sgm{4}|
\end{ffcode}

After transformation it will look like:
\begin{ffcode}
memory > flag
flag.write 0
while.
  |\m{9}|
  |\sgm{9}|
    Z'
    if.
      (eq. (flag 1)).not
      |\sgm{4}|
      TRUE
\end{ffcode}

\subsection{Final algorithm}
To sum up, we can now fully describe the algorithm itself. The first step is to make the transformations, if necessary, described in sections 3.1.1 and 3.2.1, depending on the current transformation. The second step is going through all the jumps in the program - forward and backward, and depending on the jump we carry out the required transformations: if you need to make a replacement in the while loop, containing a jump - we use transformations 3.1.3 and 3.2.3 accordingly. If you want to make a replacement for a normal object that frames the jump - use transformations 3.1.2 and 3.2.2 respectively.
